---
title: "Benchmark"
author: "Ruben Heradio (rheradio@issi.uned.es)"
date: "Date: `r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    theme: journal
bibliography: references.bib    
csl: ieee.csl    
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

# Initialization

## Importing packages and installing them if needed

```{r package_import, results='hide', warning=FALSE}
REQUIRED_PACKAGES <- 
  c("tidyverse", "gridExtra", "grid", "directlabels", "ggrepel","splines", "modelr", "broom", "psych")
lapply(
  REQUIRED_PACKAGES,
  function(pkg) {
    print(pkg)
    if (system.file(package = pkg) == "") {
      install.packages(pkg,
                       repos = "http://cran.us.r-project.org"
      )
    }
    do.call("library", list(pkg))
  }
)
```
```{r}
sessionInfo()
```

```{r global-constants-and-variables-initialization}
MODELS_PATH <- "../data"
SAT_DIST_STATS <- "samplers_stats.csv"
MODELS_EXTENSIONS <-
  c("bdd", "kus", "quicksampler", "smarch", "spur", "unigen2")
models <- list.dirs(path = MODELS_PATH, 
                    full.names = FALSE, 
                    recursive = FALSE)
SAMPLER_COLORS <- c("#E91E63", "#7EBD5E", "#3388CC", "#FA9D0D", "#666666", "#FF6E66")
SPL_MODELS <- c("axtls", "busybox", "DellSPLOT", "embtoolkit-onlybool", "fiasco", 
                "jhipster", "LargeAutomotive", "toybox", "uClibc")
FORMATTED_SPL_MODELS <- c("axTLS", "BusyBox", "DellSPLOT", "EmbToolkit", "Fiasco", 
                          "JHipster", "LargeAutomotive", "ToyBox", "uClibc")
```

# Benchmark 

The samplers were tested against a benchmark composed of 218 models encoded as Boolean formulas in all of the following formats:

+ *DIMACS*, which is the format QuickSampler, Smarch, Spur, and Unigen2 use as input. These samplers rely on SAT technology, and DIMACS is the format for *Conjunctive Normal Form* (CNF) formulas that SAT technology uses.
+ *DDDMP*, which is the format BDDSampler uses for BDDs.
+ *NNF*, which is the format KUS uses for d-DNNFs.

Nine of the 218 models represent configurable software systems. The remaining models are industrial SAT formulas (mostly modeling integrated circuits) that are typically used as a benchmark in the SAT-sampling literature [@Chakraborty15; @Achlioptas18; @Plazar19]. 

The experiments were run on an Intel(R) Core(TM) i7-6700HQ, 2.60GHz, 16GB RAM, operating Linux Ubuntu 19.10. Samplers were executed on a single thread (i.e., with no parallelization), and without considering any Boolean formulas' preprocessing, such as formula's *Minimal Independent Support* (MIS)  [@Ivrii16].

## The number of variables and clauses are highly correlated

```{r benchmark-vars-and-clauses-correlation}
# Load benchmark information
samp_data <- read_delim(str_c(MODELS_PATH, "/", SAT_DIST_STATS), delim=";")

# Transform variables and clauses into logarithmic scale
log_data <- log2(samp_data[colnames(samp_data) %in% c("vars", "clauses")])
colnames(log_data)[3:8] <- c("BDDSampler", "QuickSampler", "Smarch", "Spur", "Unigen2", "KUS")
log_data$model <- samp_data$model

# SPL models
spl_models <- log_data %>%
  filter(model %in% SPL_MODELS) %>%
  mutate(model = FORMATTED_SPL_MODELS)

# Correlation and linear regression
print("Pearson's r:")
with(log_data, cor(clauses, vars))
clauses_vars_mdl <- lm(clauses ~ vars, data=log_data)
summary(clauses_vars_mdl)

# Represent a scatter-plot

clauses_vars_data <- log_data %>% 
  add_predictions(clauses_vars_mdl, "pred_clauses") %>%
  mutate(few_clauses=ifelse(clauses < pred_clauses, TRUE, FALSE))

var_vs_clauses_plot <- ggplot(clauses_vars_data, aes(x=vars, y=clauses)) +
  geom_smooth(se=FALSE, method = "lm", color="grey", alpha=0.5) +
  scale_color_manual(values=c("#E91E63", "#3388CC", "#FA9D0D", "#666666", "#FF6E66")) +
  geom_point(aes(color=few_clauses), size=2, alpha=0.3, show.legend = FALSE) +
  geom_label_repel(data=spl_models, aes(label=model), nudge_x = -3, nudge_y=4,
                   size = 3, label.padding = 0.15, force=2000, max.time=600) +
  geom_point(data=spl_models, size=2, shape=1, show.legend = FALSE) +
  scale_x_continuous(breaks=seq(2,16, by=2)) +
  scale_x_continuous(breaks=seq(2,16, by=2)) +
  theme_minimal() +
  labs(
    x = expression(Log[2]("#Variables")),
    y = expression(Log[2]("#Clauses")),
    subtitle = "The number of variables and clauses are \nhighly correlated (Pearsons' r=0.97)"
  ) 

# Represent a histogram

quantiles <- log2(quantile(samp_data$vars, c(0.075, 0.925)))
var_plot <- ggplot(clauses_vars_data, aes(x=vars)) +
  geom_rect(xmin=quantiles[1], xmax=quantiles[2], ymin=0, ymax=Inf, fill="#FA9D0D", alpha=0.02) +
  geom_histogram(bins=30, col="#666666", alpha=0.75, size=0.1) +
  scale_x_continuous(breaks=seq(2,16, by=2)) +
  scale_y_continuous(breaks=seq(0,40, by=20)) +
  theme_minimal() +
  labs(
    x = expression(Log[2]("#Variables")),
    y = "#Models",
    subtitle = "85% of the models have between 52 and 928 \nvariables (5.70 and 9.86 in logarithmic scale)"
  ) 

complete_plot <- grid.arrange(
  var_plot,
  var_vs_clauses_plot,
  ncol = 1,
  nrow = 2,
  heights = c(1,2)
)

cairo_pdf("clauses_vs_varss.pdf", width = 4, height = 4.5)
complete_plot
dev.off()
```

## Detailed information

The following table summarizes the benchmark.

```{r benchmark-in-detail}
samp_data_detailed <- samp_data %>%
  select(model, vars, clauses) %>%
  arrange(vars)
knitr::kable(samp_data_detailed)
```

# References

