---
title: "Q3: SFpC's scalability"
author: "Ruben Heradio (rheradio@issi.uned.es)"
date: "Date: `r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    theme: journal
bibliography: references.bib    
csl: ieee.csl     
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```
# Summary

The distribution of SAT-solutions was computed with the [algorithm](https://github.com/rheradio/VMStatAnal) we presented in [@Heradio19], which takes the BDD encoding of a model as input. It is worth remarking that the model which needed the longest time was s1196a_3_2, which is an industrial SAT formula. This illustrates the dependency that BDDs have on variable ordering heuristics. Whereas this model has a medium-size CNF formula (690 variables and 1,805 clauses), the BDD we synthesized was huge (2,284,697 nodes). In contrast, for LargeAutomotive (17,365 variables and 32,1897) a more reduced BDD was obtained (30,432 nodes), and hence computing its SAT-solution distribution just took 2.74 minutes. 

# Initialization

## Importing packages and installing them if needed

```{r package_import, results='hide', warning=FALSE}
REQUIRED_PACKAGES <- 
  c("tidyverse", "gridExtra", "grid", "directlabels", "ggrepel","splines", "modelr", "scales", "kableExtra")
lapply(
  REQUIRED_PACKAGES,
  function(pkg) {
    print(pkg)
    if (system.file(package = pkg) == "") {
      install.packages(pkg,
                       repos = "http://cran.us.r-project.org"
      )
    }
    do.call("library", list(pkg))
  }
)
```
```{r}
sessionInfo()
```

```{r global-constants-and-variables-initialization}
MODELS_PATH <- "../data"
SAT_DIST_STATS <- "samplers_stats.csv"
VAR_PROB_STATS <- "samplers_stats_varprob.csv"
SAT_DIST_POPSTATS <- "population_desc_stats.csv"
VAR_PROB_POPSTATS <- "population_desc_stats_varpob.csv"
MODELS_EXTENSIONS <-
  c("bdd", "kus", "quicksampler", "smarch", "spur", "unigen2")
models <- list.dirs(path = MODELS_PATH, 
                    full.names = FALSE, 
                    recursive = FALSE)
SAMPLER_COLORS <- c("#E91E63", "#7EBD5E", "#3388CC", "#FA9D0D", "#666666", "#FF6E66")
SPL_MODELS <- c("axtls", "busybox", "DellSPLOT", "embtoolkit-onlybool", "fiasco", 
                "jhipster", "LargeAutomotive", "toybox", "uClibc")
FORMATTED_SPL_MODELS <- c("axTLS", "BusyBox", "DellSPLOT", "EmbToolkit", "Fiasco", 
                          "JHipster", "LargeAutomotive", "ToyBox", "uClibc")
```

# Time to compute the distribution of the SAT-solutions

```{r sat-dist-computation}
benchmark <- read_delim(
  file = str_c(MODELS_PATH, "/", SAT_DIST_POPSTATS),
  delim = ";"
)
benchmark$vars <- log2(benchmark$vars)
benchmark$sat_dist_time <- log2(benchmark$sat_dist_time)
sat_dist_mdl <- MASS::rlm(sat_dist_time ~ ns(vars,10), data=benchmark)
benchmark <- add_predictions(benchmark, sat_dist_mdl)

spl_models <- benchmark %>%
  filter(model %in% SPL_MODELS) %>%
  mutate(model = FORMATTED_SPL_MODELS)

ggplot(benchmark, aes(x=vars, y=sat_dist_time)) +
  geom_point(color="#E91E63", size=2, alpha=0.5) +
  geom_line(aes(y=pred), color="#E91E63", alpha=0.3, size=2) +
  geom_label_repel(data=spl_models, aes(label=model), nudge_x = -3, nudge_y=4,
                   size = 3, label.padding = 0.15, force=2000, max.time=600) +
  geom_point(data=spl_models, size=2, shape=1, show.legend = FALSE) +
  scale_x_continuous(breaks=seq(2,16, by=2)) +
  scale_y_continuous(breaks=seq(-10,10, by=2)) +
  theme_minimal() +
  labs(
    x = expression(Log[2]("#Variables")),
    y = expression(Log[2]("Time in seconds"))
  ) 
ggsave("sat_dist_time.pdf", width = 5, height = 2.75)  
  
```

## Detailed SAT-solution distribution computation time

```{r detailed-sat-dist-time}
benchmark_table <- tibble(
  model=benchmark$model, 
  vars=round(2**benchmark$vars, 0),
  clauses = benchmark$clauses,
  time=2**benchmark$sat_dist_time
)
knitr::kable(
  benchmark_table %>% 
  arrange(model) 
)
```


```{r eval=FALSE, echo=FALSE}
# Code to generate the paper latex table
numbers_with_comma <- format_format(big.mark = ",", decimal.mark = ".", scientific = FALSE)
spl_models <- spl_models %>%
  arrange(vars) %>%
  mutate(sat_dist_time = numbers_with_comma(round(2**sat_dist_time, 2))) %>%
  select(model, sat_dist_time)
kbl(spl_models, format="latex", align="l", booktabs =TRUE)
```

# Comparing the Selected Features per Configuration (SFpC) test with the Feature  Probability (FP) test

```{r read-experimental-data}
sat_dist_stats <- read_delim(
  file = str_c(MODELS_PATH, "/", SAT_DIST_STATS),
  delim = ";",
  col_names = TRUE,
  col_types = cols(
    col_character(),
    col_integer(),
    col_integer(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double()
  )
) %>% arrange(model)

var_prob_stats <- read_delim(
  file = str_c(MODELS_PATH, "/", VAR_PROB_STATS),
  delim = ";",
  col_names = TRUE,
  col_types = cols(
    col_character(),
    col_integer(),
    col_integer(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double(),
    col_double()
  )
) %>% arrange(model)
# Transform vars and sample size to a logarithmic scale
sat_dist_stats$vars <- log2(sat_dist_stats$vars)
sat_dist_stats$sample_size <- log2(sat_dist_stats$sample_size)
var_prob_stats$vars <- log2(var_prob_stats$vars)
var_prob_stats$sample_size <- log2(var_prob_stats$sample_size)
```

```{r}
# Get regression curves for plotting
sat_dist_mdl <- MASS::rlm(sample_size ~ ns(vars,20), data=sat_dist_stats)
var_prob_mdl <- MASS::rlm(sample_size ~ ns(vars,20), data=var_prob_stats)
sat_dist_stats <- add_predictions(sat_dist_stats, sat_dist_mdl)
var_prob_stats <- add_predictions(var_prob_stats, var_prob_mdl)

# Reformat data for plotting
sample_size_comparison <- tibble(
  model = c(sat_dist_stats$model, 
            var_prob_stats$model), 
  sample_size = c(sat_dist_stats$sample_size, 
                  var_prob_stats$sample_size),
  vars = c(sat_dist_stats$vars, 
           var_prob_stats$vars),
  pred = c(sat_dist_stats$pred, var_prob_stats$pred),
  method = factor(c(rep("SFpC test", nrow(sat_dist_stats)), 
                        rep("FP test", nrow(var_prob_stats))),
                     levels = c("FP test", "SFpC test"))
)

# Get tibbles for tagging configuration models
spl_models_sat_dist <- sat_dist_stats %>%
  filter(model %in% SPL_MODELS) %>%
  mutate(model = FORMATTED_SPL_MODELS)
spl_models_var_prob <- var_prob_stats %>%
  filter(model %in% SPL_MODELS) %>%
  mutate(model = FORMATTED_SPL_MODELS)

# Plot a comparison between the SAT-solution distribution goodness-of-fit test and the
# one based on examining the variable probabilities
ggplot(sample_size_comparison, aes(x=vars, y=sample_size)) +
  scale_color_manual(values=c("#E91E63", "#3388CC", "#FA9D0D", "#666666", "#FF6E66")) +
  geom_point(size=2, alpha=0.3, aes(color=method)) +
  geom_line(aes(y=pred, color=method), size=2, alpha=0.4) +
  geom_point(data=spl_models_var_prob, size=2, shape=1, color="#E91E63") +
  geom_label_repel(data=spl_models_var_prob, aes(label=model), nudge_x = -0.5, nudge_y=1,
                   size = 2, label.padding = 0.15, force=1, max.time=600,
                   color="#E91E63") +
  geom_point(data=spl_models_sat_dist, size=2, shape=1, color="#3388CC") +
  geom_label_repel(data=spl_models_sat_dist, aes(label=model), nudge_x = 0.5, nudge_y=-1,
                   size = 2, label.padding = 0.15, force=1, max.time=600,
                   color="#3388CC") +
  scale_x_continuous(breaks=seq(2,16, by=2)) +
  scale_y_continuous(breaks=seq(10,17, by=1)) +
  theme_minimal() +
  theme(legend.position="bottom") +
  labs(
    x = expression(Log[2]("#Variables")),
    y = expression(Log[2]("Sample size")),
    subtitle = "The SFpC test requires smaller samples than the FP test",
    color = ""
  )  
ggsave("sat_dist_vs_var_prob.pdf", width = 4.5, height = 3.5)
```

## Detailed sample size comparison

```{r result-tibble-storage}
sample_size_comparison <- tibble(
  model=sample_size_comparison$model, 
  vars=2**sample_size_comparison$vars, 
  sample_size=2**sample_size_comparison$sample_size, 
  goodness_of_fit_test=sample_size_comparison$method
)
knitr::kable(
  sample_size_comparison %>% 
  arrange(model) 
)
```


```{r eval=FALSE, echo=FALSE}
# Code to generate the paper latex table
numbers_with_comma <- format_format(big.mark = ",", decimal.mark = ".", scientific = FALSE)
spl_models <- inner_join(spl_models_sat_dist, spl_models_var_prob, by="model") %>%
  select(model, vars.x, sample_size.x, sample_size.y) %>%
  mutate(satdist_sample_size = numbers_with_comma(round(2**sample_size.x, 0)),
         varprob_sample_size = numbers_with_comma(round(2**sample_size.y, 0))) %>%
  arrange(vars.x) %>%
  select(model, satdist_sample_size, varprob_sample_size)
kbl(spl_models, format="latex", align="l", booktabs =TRUE)
```

# References

